---
title: "Data Management Project"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## **Data Analytics Overview**

### WHAT IS DATA SCIENCE?

  1. Data Science allows us to extract knowledge from databases and open data portals.  
  2. Lets people leverage large amounts of data and computing power to tackle complex questions.  
  3. Patterns can arise which could not have been uncovered otherwise.  
  4. These discoveries can lead to powerful results.


### METHODOLOGY

Data Science Methodology is adapted from the Cross-Industry Standard Practice for Data Mining (CRISP-DM) and includes:

```{r fig1, fig.align="center", echo=FALSE, fig.cap="**CRISP-DM Methodology**", out.width = '70%',}
knitr::include_graphics("https://drive.google.com/uc?id=1iYx4-h079unPjL2M4oh29jj86FJyTqDp")
```
                      
1. Problem Understanding Phase:
    + Teams often work hard to solve a problem, only to find they solved the wrong problem.
    + Enunciate project 6objectives
    + Translate objectives into a problem that can be solved using data science.

2. Data Preparation Phase: 
    + Raw data from data repositories needs to be cleaned.  
    + Probably the most labor-intensive phase:
        + Identifying outliers.
        + Transforming and standardizing the data.
        + Reclassifying categorical variables.
        + Binning numerical variables.
        + Adding an index field.
        
3. Exploratory Data Analysis Phase:  
    + Begin to explore and learn basic information.  
    + Not the time for complex algorithms. 
    + Exploring univariate relationships 
    + Exploring multivariate relationships among the variables.
    + Binning based on predictive value 
    + Deriving new variables
        
4. Setup Phase:  
    + Nearly ready to model  
    + Need to take care of Cross-validation necessary to avoid data dredging.Data partitions need to be evaluated.
    + Need to take care of Balancing the data to enhances the ability of certain algorithms

5. Modeling Phase:  
    + Apply algorithms to uncover relationships hidden in the data.  
    + The heart of data scientific investigation:
        + Selecting and implementing appropriate algorithms: Inappropriate techniques will lead to inaccurate results
        + Making sure that our models outperform the baseline models.  
        + Fine-tuning your model algorithms: What should be our cutoff point to maximize profits?  
        + Need to spend time fine-tuning models before arriving at the optimal solution.
        
6. Evaluation Phase:
    + Anyone can make predictions.
    + How do the predictions perform against real data?
    + Assess how models are doing, and whether we need to go back and try to improve.
    + Your models need to be evaluated against the baseline performance measures from the Setup Phase.  
    + You need to determine whether your models are actually solving the problem at hand.  
    + Apply error costs intrinsic to the data, because data-driven cost evaluation is the best way to model the actual costs involved.
    
7. Deployment Phase:
    + Report to management on your best models
    + Work with management to adapt your models for deployment.
    + Writing a report: a simple example
    + Stay involved with the project!  
    + Participate in the meetings and processes involved in model deployment, so that they stay focused on the problem at hand.



```{r quiz, echo=FALSE}
quiz(
  question("Data Science is:",
    answer("adaptive"),
    answer("iterative"),
    answer("phased approach to analysis of data"),
    answer("that uncover optimal models"),
    answer("all of the above", correct = TRUE),
    allow_retry = TRUE
  )
)
```

## **Data Science Task**

The most common data science tasks are the following:

1. Description: To describe patterns and trends  
    + COVID 19 Dashboards including information on number of cases, admission to hospital, geographical information down to national, regional  and postcode level.
    + Demographic information about the age and sex of patients/cases.
    + Current [COVID-19 Trends](https://covid19.who.int/)
  
2. Estimation: The approximation of the value of a numeric target variable using predictor variables.  

3. Classification
4. Clustering
5. Prediction
6. Association

## **Introduction to R**

This tutorial will help you set up your computer to use R. It is for you if you need to:

* Install R on your computer
* Install the RStudio IDE
* Install the `tidyverse` R package or any other R packages

You can skip this tutorial if you've already done these things.

### Is this tutorial for you?

Do you need to work through the tutorial? Take the quiz below to find out.


```{r quiz1, echo = FALSE}
question("Check all that you have NOT done:",
  answer("installed R on my computer", message = "* Install R"),
  answer("installed the RStudio IDE", message = "* Install RStudio IDE"),
  answer("installed the tidyverse R package", message = "* Install Packages"),
  answer("None of the above. I've done them all.", correct = TRUE, message = "You can skip this tutorial!"),
  type = "multiple",
  incorrect = "This tutorial is here to help! To get set up read:"
)
```

## **Install R**

### How to install R

![](https://vimeo.com/203516510)

### Test your knowledge

```{r quiz2, echo=FALSE}
quiz(caption = "Quiz - Install R",
  question("Is R free to download and use?",
    answer("Yes!", correct = TRUE, message = "R is free _and_ open source, which means that anyone can read, use, and copy the source code of the R language."),
    answer("No.")
  ),
  question("How often should you update R?",
    answer("Everytime you use it", message = "This will be too often unless you use R very rarely!"),
    answer("About once a year", correct = TRUE, "A new version of R is released about once a year. Update sooner if you encounter a bug that you cannot explain."),
    answer("Never", message = "A new version of R is released about once a year. I'll assume that you are using the newest version of R, which will be the fastest version with the fewest unexpected behaviors." ),
    allow_retry = TRUE
  )
)
```

## **Install RStudio**

### How to install RStudio

RStudio is an Integrated Development Environment for R. What does that mean? Well, if you think of R as a language, which it is, you can think of RStudio as a program that helps you write and work in the language. RStudio makes programming in R much easier and I suggest that you use it!

![](https://vimeo.com/203516968)

### Test your knowledge

```{r quiz3, echo=FALSE}
quiz(caption = "Quiz - Install RStudio",
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use R.", correct = TRUE, message = "RStudio organizes your workspace and makes it easier to write, use, debug, and save R code. I highly recommend it and will rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code", message = "And thank goodness! Code provides a reproducible record of your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel."),
    answer("Another name for R", message = "R and RStudio are two separate things. R is a language, like English. Think of RStudio as a program that helps you use the language, kind of like how a word processing program helps you write in English."),
    allow_retry = TRUE
  ),
  question("Is the RStudio IDE free to download and use?",
    answer("Yes!", correct = TRUE, message = "Like R, RStudio is free and open-source. There do exist professional versions of RStudio with enhanced features, but we will not rely on those features in these tutorials."),
    answer("No.", message = "Like R, RStudio is free and open-source. There do exist professional versions of RStudio with enhanced features, but we will not rely on those features in these tutorials.")
  ),
  question("Where do you download RStudio?",
    answer("www.rstudio.com/download", correct = TRUE, message = "For these tutorials, download and install the RStudio Desktop Open Source License."),
    answer("[cloud.r-project.org](http://cloud.r-project.org)", message = "This is where you download R, not RStudio."),
    answer("www.r-project.org"),
    answer("[cran.rstudio.org](http://cran.rstudio.org)"),
    allow_retry = TRUE
  ),
  question("Do you need to install R if you already have RStudio?",
    answer("Yes.", correct = TRUE),
    answer("No.", message = "R does not come with RStudio; you need to install R separately.")
  )
)
```

## **Install Packages**

### How to install R packages

![](https://vimeo.com/203516241)

### Test your knowledge

```{r names, echo = FALSE}
quiz(caption = "Quiz - Working with Packages",
  question("What command do you use to install packages?",
    answer("`library()`", message = "We will see what `library()` does later."),
    answer("`install.packages()`", correct = TRUE),
    answer("`install_packages()`"),
    answer("There is no command. You must visit [cran.r-project.org](http://cran.r-project.org) and download packages manually.", message = "R makes it easy to download packages. You connect to the internet and then run one of the commands above."),
    allow_retry = TRUE
  ),
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R"),
    answer("Every time you restart your computer"),
    answer("Only once. Afterwards, R can find it on your hard drive as needed.", correct = TRUE),
    answer("Never, as long as you are connected to the internet.", message = "This could be true if you are using R over a cloud service. However, if you are using R locally on your own computer, you will need to install each package that you use on your computer."),
    allow_retry = TRUE
  ),
  question("What is the tidyverse?",
    answer("A collection of packages that work well together and provide tools for common data science tasks.", correct = TRUE, message = 'You can install all of the tidyverse packages at once with `install.packages("tidyverse")`.'),
    answer("A word processing program for R"),
    answer("A code style that reads like poetry"),
    answer("An alternate universe where everyone is an R programmer."),
    allow_retry = TRUE
  )
)
```


*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:


